---
title: "2024_12_04_latest_paper_draft"
author: Alejandra Quinones
output: html_document
---

```{r}
library(tidyverse)
library(ggplot2); theme_set(theme_bw())
library(pcvr)
library(ggpubr)
library(scales)
library(pcvr)
library(tidygraph)
library(ggraph)
library(geomtextpath)
library(corrplot)
library(gridGraphics)
library(grid)
library(multcompView)
library(corrr)
library(ggridges)
library(ppclust)
library(ggalluvial)

resp_traits = c("area_cm2", "height_above_reference_cm", 
                          "hue_circular_mean_degrees", "pWUE")
short_names_traits = c("area_cm2"="Plant area (cm²)","height_above_reference_cm"="Height (cm)", 
                          "hue_circular_mean_degrees"="Plant mean hue (°)", "pWUE"="pWUE")

treat_cols = c("Control"="#7CAE00", "Drought"="#00BFC4","Heat"="#F8766D", "Drought_Heat"="#C77CFF")
treat_fullname = c("Control"="Control","Drought"="Drought","Drought_Heat"="Drought & Heat", "Heat"="Heat")
treat_short = c("C"="Control","D"="Drought","DH"="Drought_Heat", "H"="Heat")

df_full = read.csv("/Users/quino070/Bellwether_RGB/2024_12_draft/2024_12_04_full_df.csv") 
```


1. Treatments affected plant size, color, and development

```{r, fig.width=15, fig.height=15}
for_varcomps = df_full %>% filter(day==13) %>% select(-c(leaf_count, pWUE, area_pixels, center_of_mass.y_converted, 
             height_above_reference_pixels))

traits_ = colnames(for_varcomps[,-c(1:6)])
  
traits =   data.frame(ps=colnames(for_varcomps) %>% .[-c(1:6)]) %>% mutate(
  p=str_replace_all(ps, "_", " ") %>%
    str_replace(., " pixels", "") %>%
    str_replace(., " none", "") %>% str_replace(., " degrees", "") %>%
    str_replace(., " cm2","") %>% str_replace(., " cm", "") %>%
    str_replace(., "[.]", " ") %>%
    ifelse(.!="pWUE", str_to_sentence(.), .)) %>% deframe()

#colnames(for_varcomps)[7:ncol(for_varcomps)] = traits[colnames(for_varcomps)[7:ncol(for_varcomps)] ]



vc = pcvr::frem(for_varcomps, des=c("treatment","genotype_name"), 
                phenotypes=traits_, returnData = T) %>% .$data

comps = c("Unexplained"="Residual", "Interaction"="Genotype x Treatment", 
          "genotype_name"="Genotype", "treatment"="Treatment")

ord_traits = vc$variance %>% mutate(Phenotypes=as.character(Phenotypes) %>% traits[.]) %>% 
  pivot_longer(cols=1:4, names_to = "Component") %>% 
  mutate(Component=factor(comps[Component], comps)) %>% 
  filter(Component == "Treatment") %>% arrange(value) %>% .$Phenotypes


var_cor = ggarrange(

vc$variance %>% mutate(Phenotypes=as.character(Phenotypes) %>% traits[.]) %>% 
  pivot_longer(cols=1:4, names_to = "Component") %>% 
  mutate(Component=factor(comps[Component], comps)) %>% 
  ggplot(aes(value, fct_relevel(Phenotypes, ord_traits), fill=Component)) + 
  geom_bar(stat="identity") +
  scale_fill_viridis_d() + 
  scale_x_continuous(labels = scales::percent, expand=c(0.1,0)) +
  scale_y_discrete(expand=c(0,0)) + 
  theme_classic() + theme(axis.line = element_blank(), 
                          axis.ticks.y = element_blank(),
                          legend.position = "bottom",
                          axis.text.y=element_blank(),
                          axis.title.x=element_text(size=15),
                          plot.margin = unit(c(6,0.5,0.65,0.25),"cm")) + 
  labs(x="Variance explained", y=NULL) 
,

vc$cor %>% mutate_at(c("Var2", "Var1"), ~{as.character(.x) %>% traits[.] 
  }) %>%
  mutate(Var1=factor(Var1, ord_traits), Var2=factor(Var2, rev(ord_traits))) %>% 
  ggplot(aes(Var2, Var1, fill=value)) + geom_point(shape=21, size=10, stroke=0) + 
  theme_pubr() + scale_fill_distiller(palette = "RdBu", direction = +1) + 
  theme(axis.line = element_blank(), axis.ticks=element_blank(), 
        axis.text.x=element_text(angle=45, hjust=0, size=15), 
        legend.position = "bottom", plot.margin = unit(c(0.25,10,0.6,0),"lines"),
        legend.title.position = "top", 
        legend.title = element_text(hjust=0.5, face="italic"),
        axis.text.y = element_text(size=15, hjust=0.5)
        ) + scale_x_discrete(position="top", expand=c(0.1,0)) + labs(fill="r", x=NULL, y=NULL)

, widths = c(0.5,1), heights = c(1,0.8)

)

#ggsave("/Users/quino070/Bellwether_RGB/2025_paper/var_comps.pdf", width = 20)
```

```{r, fig.height=20}

treat_dayBU = imap(1:13, ~{
  temp_vc = df_full %>% filter(day==.x) %>% select(-c(leaf_count, pWUE, area_pixels, center_of_mass.y_converted, 
             height_above_reference_pixels))

  pcvr::frem(temp_vc, des=c("treatment","genotype_name"), 
                phenotypes=traits_, returnData = T, cor = F) %>% .$data %>% 
    mutate(Phenotypes=as.character(Phenotypes) %>% traits[.]) %>% 
  pivot_longer(cols=1:4, names_to = "Component") %>% 
  mutate(Component=factor(comps[Component], comps), day=.x) 
    
}) %>% do.call("rbind",.) 


treat_day = treat_dayBU %>% filter(Component=="Treatment") %>% 
  ggplot(aes(as.factor(day), fct_relevel(Phenotypes,ord_traits), fill=value)) + geom_tile() + 
  theme_pubclean() + scale_fill_viridis_c(name="Variance explained\n      by treatment", 
                                           labels=scales::percent, 
                                          option = "F", direction = -1) +
  labs(x="Day of experiment", y=NULL) + scale_x_discrete(position = "top") + 
  theme(legend.position = "bottom", legend.key.width = unit(1.5, "cm"))




ggarrange(var_cor, ggplot() + theme_void(), ggarrange(treat_day, ggplot() + theme_void(), 
                             widths=c(0.3,0.6)), ncol=1, heights = c(1,0.05,0.5))

#ggsave("/Users/quino070/Bellwether_RGB/2025_paper/var_comps.png", width=20, height = 20)
```


Calculating %similarity in general trends
```{r, fig.height=12}

# long = df_full %>% dplyr::select(1:6, resp_traits[-4]) %>%
# pivot_longer(cols = resp_traits[-4], names_to = "trait")

long = df_full %>% dplyr::select(1:6, resp_traits) %>%
pivot_longer(cols = resp_traits, names_to = "trait")

gen_trend = 
  #general trends dataset
  left_join(
  long %>% group_by(treatment, day, trait)  %>% 
  summarize(mean_t=mean(value, na.rm = T), 
            sd = sd(value, na.rm=T),
            min=mean_t - 1.96*(sd/sqrt(47)) , max=mean_t + 1.96*(sd/sqrt(47))) %>% 
  ungroup(),
  #stat test to get first instance of significant change from control
  imap(resp_traits, ~{
    t=.x
    days = if(.x == "pWUE"){
      2:13}
    else 1:13
    imap(days, ~{
      aov(value ~ treatment + genotype_name + treatment:genotype_name, data = long %>% filter(trait == t & day==.x)) %>% 
        TukeyHSD() %>% .$treatment %>% data.frame() %>% 
        rownames_to_column("n") %>% rowwise() %>% 
        mutate(trait=t, day=.x,
               treatment=str_split(n, "-") %>% .[[1]] %>% .[1],
               treatment2=str_split(n, "-") %>% .[[1]] %>% .[2]) %>% ungroup() %>% 
        filter(treatment2=="Control") %>% select(treatment, day, trait,2:6)
      }) %>% do.call("rbind",.) 
    })  %>%  do.call("rbind",.) %>% mutate(sig=ifelse(p.adj<0.05, "*", NA))
  )
```


```{r}
gen_trend = gen_trend %>% left_join(., 

gen_trend %>% group_by(trait) %>% 
  mutate(terminal_C = mean_t[treatment=="Control" & day==13],
         similarity=mean_t/terminal_C,
         sim_max = max/max[treatment=="Control" & day==13],
         sim_max = min/min[treatment=="Control" & day==13]
         
         ) %>% 
  filter(day==13 & treatment!="Control") 
) #%>% filter(trait %in% c("area_cm2", "height_above_reference_cm")) 

```

```{r, fig.height=10, fig.width=10}
gt_plots = imap(c("area_cm2", "height_above_reference_cm", "hue_circular_mean_degrees","pWUE"), ~{
  gen_trend %>% filter(trait==.x) %>% 
    ggplot() + 
  geom_line(aes(day, mean_t, group=treatment, color=treatment)
            #, linewidth=2
            ) + 
  geom_ribbon(aes(day, mean_t, group=treatment, ymax=max, ymin=min,fill=treatment), alpha=0.25) + 
  geom_point(data= gen_trend %>% filter(trait==.x & sig=="*"), 
             aes(day, mean_t, group=treatment, color=treatment)
             #, size=10
             ) + 
  labs(y=NULL, fill="Treatment", color="Treatment", x="Days of treatment") +
  
  geom_pointrange(data= gen_trend %>% filter(trait==.x & day==13 & treatment!="Control") %>% 
                    mutate(day=
                             #c(13.25, 13.5, rep(13,2), 13.5, 13.25)
                           ifelse(trait %in% c("area_cm2", "hue_circular_mean_degrees"), c(13.25, 13.50, 13.00),
                                  ifelse(trait=="pWUE",c(13.50,13.25,13.00),c(13.00, 13.50, 13.25)))
                             ), 
                  aes(day, mean_t, ymax=mean_t, ymin=-Inf,color=treatment), 
                  #size=5, 
                  shape=17, 
                  #linewidth=3, 
                  show.legend = F) +
  scale_fill_manual(values=treat_cols, labels=treat_fullname) + 
  scale_color_manual(values=treat_cols, labels=treat_fullname) +
  
  geom_text(data=. %>% filter(trait==.x & day==13 & treatment!="Control") %>% 
              mutate(day=
                       # c(13.25, 13.5, rep(13,2), 13.5, 13.25)
                       ifelse(trait %in% c("area_cm2", "hue_circular_mean_degrees"), c(13.25, 13.50, 13.00)+0.25,
                                  ifelse(trait=="pWUE",c(14,13.5,13.25),c(13.00, 13.50, 13.25)+0.25))
                       #rep(14.5,3)
                     
                     ), 
            aes(day, label=round(similarity*100,0) %>% paste0(.,"%"),
                y=mean_t, color=treatment), #size=20, #fontface="bold",
            hjust=-0.1, show.legend = F) + 
  scale_x_continuous(expand=c(0,0), 
                     limits = c(1,15.5), breaks = c(1,7,13)) +
  theme_pubr() +
  scale_y_continuous(name=short_names_traits[.x]#, expand=c(0,0)
                     ) + 
  theme(legend.position="top", legend.title = element_blank(), 
        legend.key.width = unit(1.25,"cm"),
        legend.text = element_text(#size=75,
                                   margin = margin(l=0.5,r = 1, unit = "cm"))
        )
}) 



gt_plots[[3]] = gt_plots[[3]] + scale_x_continuous(expand=c(0,0), 
                     limits = c(0,15.5), breaks = c(1,7,13)) + 
  scale_y_continuous(expand=c(0,0)) +
  
  ggnewscale::new_scale_fill() +
  geom_raster(data=
              #data.frame(x=rep(1:13, 56), y=rep(40:95, 13))
              data.frame(x=rep(c(0.5), each=27),y=rep(66:92,1))
            ,
            aes(x,y, fill=y), show.legend = F) +
  scale_fill_gradient2(high="#98DF65", mid="#CADE64",
                       low="#DEBF64", midpoint = 70) 


ggarrange(plotlist = gt_plots, common.legend = T)

# ggsave("/Users/quino070/Bellwether_RGB/2025_paper/gen_trends.pdf",
#        width = 7.5, height=8, units = "in", limitsize = F)
```








Getting similarity metric for every day. Using the formula:
$$
%similarity = \frac{mean_treatment}{mean_Control_day13}
$$

```{r message=FALSE, warning=FALSE}
summary_df = df_full %>% dplyr::select(1:6, resp_traits) %>% 
  pivot_longer(cols = resp_traits, names_to = "trait") %>% 
  group_by(genotype_name, treatment, day, trait) %>% 
  summarize(mean=mean(value, na.rm=T), 
            #adding variance calculation and a count of how many replicates for genotype:treatment to make my life easier later
            var=var(value, na.rm=T), n=n()) %>% 
  group_by(genotype_name, trait) %>% 
    mutate(terminal_C = mean[treatment=="Control" & day==13],
      #similarity metric is being calculated in relation to the value in control at the last day of the experiment
      similarity = mean/mean[treatment=="Control"])
```


```{r, fig.width=20, fig.height=30}
# pdf("/Users/quino070/Bellwether_RGB/2025_paper/similarity.pdf", width = 10, height = 12)

ord = summary_df %>% filter(trait=="area_cm2" & treatment=="Drought" & day==13) %>% 
  arrange(-similarity) %>% .$genotype_name


imap(resp_traits, ~{
  df = summary_df %>% filter(trait==.x & treatment!="Control") %>% 
    mutate(similarity=ifelse(similarity>1, 1, similarity)) 
    fill_brks=c(0,0.25,0.50,0.75,1)
      fill_limits = c(0,1)
  
  if(.x=="pWUE"){brks= 2:13} else {brks=1:13}
  
  plt = df %>% filter(!is.na(similarity)) %>% 
  ggplot(aes(as.factor(day), fct_relevel(genotype_name, ord), fill=similarity)) + geom_tile() +
  facet_wrap(~treatment, nrow=1, labeller = as_labeller(~treat_fullname[.x])) +
    scale_x_discrete(breaks=brks) +
    labs(x="Day of Experiment", y=NULL, fill="Similarity") + 
    #geom_text(aes(label = round(similarity,2) *100), size=7, show.legend = F) +
    ggtitle(short_names_traits[.x]) + 
    theme(strip.text.x = element_text(size=15),
          axis.text=element_text(size=12))
  
    plt + scale_fill_distiller(palette = "Spectral", type = "seq",
                         labels=scales::percent_format(),
                         direction = +1, 
                         breaks=fill_brks, 
                         limits=fill_limits,
                         guide = guide_colorbar(barheight=20)
                         )   


  #ggsave(paste0("/Users/quino070/Bellwether_RGB/2025_paper/similarity_", .x, ".png"), width=10, height=12)
  
  })

# dev.off()


```


Sources
```{r}
library(gtExtras)

a = df_full %>% select(genotype_name, subpopulation) %>% unique() %>% 
  full_join(.,read.csv("/Users/quino070/Downloads/bellwether_source.csv")) %>% 
  #modify label for plotting purposes
  mutate(subpopulation=ifelse(
    subpopulation=="Temporate/Tropical mix", "Temporate/ \n Tropical mix", 
    ifelse(subpopulation == "Stiff-stalk", "Stiff stalk",
           ifelse(subpopulation == "Non-stiff-stalk", "Non-stiff stalk", 
                  subpopulation))
    )
    )

source = a %>% 
  #mutate(source = factor(source, c("NAM", "WiDiv", "Specialized"))) %>% arrange(source, subpopulation) %>%  
  rename("Genotype"='genotype_name', 
             "Subpopulation"="subpopulation", "Main Source"="source") %>% 
  mutate(Subpopulation=ifelse(Subpopulation=="Temporate/ \n Tropical mix", 
                              "Temperate/Tropical mix", Subpopulation),
         `Main Source` = ifelse(`Main Source`=="NAM", "NAM founder", `Main Source`) %>% 
           factor(., c("NAM founder", "WiDiv", "Specialized"))) %>% 
  arrange(`Main Source`, Subpopulation) 
  
source = gt(source) %>% tab_style(style = cell_text(weight = "bold"), 
                         locations = cells_column_labels())

#gtsave(source, filename = "/Users/quino070/Bellwether_RGB/2025_paper/germplasm.pdf")

```


#Subpopulation differences
Testing subpopulation as a factor


```{r, fig.width=10, fig.height=10}
library(betareg)
library(lmtest)
library(ggtext)


spds = 
df_full %>% select(genotype_name, subpopulation) %>% unique() %>% 
  left_join(summary_df,.) %>% 
  filter(day==13 & treatment!="Control" &
           subpopulation %in% c("Tropical", "Stiff-stalk", "Non-stiff-stalk"))

subpop_resp_plots = spds %>% 
  ggplot(aes(similarity, subpopulation)) + 
  geom_density_ridges(aes(fill=subpopulation),
                      alpha=0.5, show.legend = F, rel_min_height = 0.01,
                      scale=0.6,
    jittered_points = TRUE, point_shape=21,
    point_alpha=1,
    point_size=2,
    position = position_points_jitter(width = 0.05, height = 0)
    ) + 
  facet_grid(treatment~trait, scales = "free", 
             labeller = labeller(.cols=~short_names_traits[.x],
                                 .rows=~treat_fullname[.x]) ) + 
  scale_fill_brewer(palette = "Spectral") + 
  scale_color_brewer(palette = "Spectral") + 
  scale_x_continuous(labels = percent) +
  labs(y="Subpopulation", x="Similarity to control")

  


kwt_subpop = imap(names(treat_fullname)[-1], ~{
  tr=.x
  imap(resp_traits, ~{
    tdf = spds %>% filter(treatment==tr & trait==.x)
    null = betareg(similarity ~ 1, data= tdf)
    mod = betareg(similarity ~ subpopulation, data=tdf)
    data.frame(
               treatment=tr,trait=.x,
               chi_squared=round(as.data.frame.list(lrtest(null,mod))[2,4],2), 
               pval=round(as.data.frame.list(lrtest(null,mod))[2,5],3))
  }) %>%  do.call("rbind",.)
}) %>% do.call("rbind",.) %>% 
  mutate(label=paste0("X² = ",round(chi_squared,2),"; p = ", round(pval,2)))


subpop_resp_plots + geom_text(data=kwt_subpop, 
                              aes(x=Inf, y=Inf, label=label), 
                              #vjust=1.5,hjust=1.5
                              vjust=2, hjust=1.5
                              ) +
  theme(strip.background = element_rect(fill=NA)) +
  scale_y_discrete(labels=c("Tropical"="Tropical", "Stiff-stalk"= "Stiff stalk", 
                            "Non-stiff-stalk"= "Non-stiff stalk"),
                   expand=expansion(mult=c(0.05,0.4)))

#ggsave("/Users/quino070/Bellwether_RGB/2025_paper/similarity_by_subpop.pdf")



```


```{r}
library(emmeans)

betareg(similarity ~ subpopulation, data=
          spds %>% filter(treatment=="Drought" & trait=="area_cm2")) %>% 
  emmeans(., pairwise ~ subpopulation)
  

betareg(similarity ~ subpopulation, data=
          spds %>% filter(treatment=="Drought" & trait=="height_above_reference_cm")) %>% 
  emmeans(., pairwise ~ subpopulation)

```

Leaf count analysis by genotype :)


```{r message=FALSE, fig.height=10}
library(dunn.test)

leaf_counts_by_gen = df_full %>% select(1:6, leaf_count) %>% filter(day==13) %>% group_by(genotype_name) %>% 
  group_map(.keep = T,~{
   dt = dunn.test(.x$leaf_count, .x$treatment, method="bonferroni")
   dtr = data.frame(genotype_name=unique(.x$genotype_name),treatment=dt$comparisons, pval=dt$P.adjusted) %>%
     filter(startsWith(treatment,"Control")) %>%
     mutate(sig=ifelse(pval<0.05, "*", ""), treatment=sub('.*- ', '', treatment))
  }) %>% do.call("rbind",.) %>% 
  full_join(., df_full %>% filter(day==13) %>% 
              group_by(genotype_name, treatment) %>% 
              summarize(leaf_count=median(leaf_count, na.rm=T) %>% as.integer()))

lcor = leaf_counts_by_gen %>% filter(treatment=="Drought") %>% 
  arrange(-leaf_count) %>% .$genotype_name

ggplot(leaf_counts_by_gen, 
       aes(treatment, fct_relevel(genotype_name, ord), 
           fill=as.factor(leaf_count), label=sig)) + 
  geom_tile(color="darkgrey") + geom_text() + 
  scale_x_discrete(expand=c(0,0), labels=~treat_fullname[.x]) +
  scale_fill_brewer(palette = "Greens", direction = +1) +
  labs(fill="Leaf count", y=NULL, x=NULL)  

#ggsave("/Users/quino070/Bellwether_RGB/2025_paper/leaf_counts_dunntest.pdf", height=12, width = 6)
```

# Multiplicative risk model:

First, calculate expected additive effect of D + H as

$$
ExpectedDH = D_{red} + H_{red} =  (D_{red} + H_{red}) - (D_{red} * H_{red})
$$

Then, calculate variance of the expected additive effect:
$$
var_{Expected_{D+H}} = var(D_{red}) * (1-H_{red})^2 + var(H_{red})*(1-D_{red})^2 
$$
Now use the expected variance to calculate a 95% confidence interval:
$$
95\%CI_{D+H} = 2.228*\sqrt{var_{Expected_{D+H}}/n}
$$
If confidence intervals overlap = additive
If Observed - Expected > 0, and CIs do not overlap = synergistic
If Observed - Expected < 0,and CI do not overlap = antagonistic


First, running anova + pairwise comparisons to get the effect of each treatment for all genotypes at each trait

```{r}
comp_means_df = imap(resp_traits, ~{
  #this first imap loop goes through all the traits
  tr = .x
  all_gens = imap(unique(df_full$genotype_name), ~{
    gen = .x
    #this second one loops through all the genotypes
    sub_df = df_full %>% filter(day==13 & genotype_name==gen)
    an = aov(eval(parse(text=tr)) ~ treatment, sub_df)
    #This dataframe contains the reductions caused by each treatment, compared to control
    tuk = 
      TukeyHSD(an) %>% .$treatment %>% as.data.frame() %>% 
      rownames_to_column("contrast") %>% rowwise() %>% 
      mutate(
        treatment=str_split(contrast, "-") %>% unlist() %>% .[[1]],
             t2=str_split(contrast, "-") %>% unlist() %>% .[[2]]
        ) %>% 
      filter(t2 =="Control") %>% rename("reduction"="diff", "pval"="p adj") %>% 
      dplyr::select(-c(contrast, t2)) %>% ungroup()
    
    sig_test = cbind(trait=tr, genotype_name=gen, tuk)
    
    return(sig_test)
  }) %>% do.call("rbind",.)
}) %>% do.call("rbind",.) %>% 
  #merge with summary dataset
  left_join(summary_df %>% filter(day==13),.) %>% group_by(genotype_name, trait) %>% 
  mutate(
    #calculating the effect as the reduction caused by the treatment over the value in control
    effect = abs(reduction/terminal_C)
    )
```


Now, I will be calculating a variance based on the percent reduction calculated for each genotype, so both the observed and expected reductions will be made off 47 datapoints.





Running model by genotype
```{r}
effects_by_gen = comp_means_df %>% 
  mutate(
  #Step 1: calculate effect (mean reduction) and variance of effect (variance of reduction)
  #first part was already done, now calculate variance of effect - using the delta method as per Darling et al 2010.
  var_effect = var*(1/terminal_C)^2 + var[treatment=="Control"]*
    (-mean/terminal_C^2)^2
  ) %>% filter(treatment != "Control")

effects_by_gen = effects_by_gen %>% 
  summarize(
    #Step 2: calculate additive null expectation
    #calculate expected DH effect
    expectedDH_effect = 
      effect[treatment=="Drought"] + effect[treatment=="Heat"] -
    (effect[treatment=="Drought"]*effect[treatment=="Heat"]),
    #calculate expected DH effect variance - once again following methods by Darling et al 2010
    expectedDH_var_effect =
      var_effect[treatment=="Drought"] * (1-effect[treatment=="Heat"])^2 +
      var_effect[treatment=="Heat"]*(1-effect[treatment=="Drought"])^2,
    #sum of observations in each subset to calculate CI later
    n = n[treatment=="Drought"] + n[treatment=="Heat"]) %>% 
  pivot_longer(cols=3:4, names_to=c("treatment", ".value"), names_sep="_") %>%
  rename("var_effect"="var") %>% 
  rbind(effects_by_gen %>% select(1:2, 4, effect, var_effect, n), .) %>% 
  mutate(CI=2.228*sqrt(var_effect/n), max=effect+CI,min=effect-CI)

```

Hypothesis test by genotype
```{r, fig.height=15}
hyp_test_by_gen = effects_by_gen %>% filter(treatment%in%c("Drought_Heat", "expectedDH")) %>% 
  summarize(hypothesis_effect=
              #Hypothesis: Observed combined effect and expected combined effect are similar
              effect[treatment=="Drought_Heat"]-effect[treatment=="expectedDH"],
            #Getting confidence interval of observed-expected
            min=min[treatment=="Drought_Heat"]-min[treatment=="expectedDH"],
            max=max[treatment=="Drought_Heat"]-max[treatment=="expectedDH"],
            #Color-coding based on results
            col=ifelse(
              #if CIs overlap, effect = additive, color=white
              min*max<=0, "Additive", 
              #if observed < expected = antagonistic
              ifelse(max<0, "Antagonistic", 
              #if observed > expected = synergistic
              "Synergistic"))) %>% mutate(col=fct_relevel(col,c("Antagonistic","Additive", "Synergistic")))

```

```{r}
hyp_test_by_gen %>% 
  ggplot(aes(trait,hypothesis_effect)) + 
  geom_violin(draw_quantiles = 0.5#, linewidth=2
              ) + 
  # geom_pointrange(aes(group=genotype_name,fill=col, ymin=min, ymax=max), 
  #                 shape=23, linetype="dashed", size=2, alpha=0.8,
  #                 position=position_jitter()
  #                 ) + 
  ggbeeswarm::geom_quasirandom(
    #shape=23, 
    size=3,
    aes(fill=col,
                shape=col)) +
  
  theme_pubr() + 
   scale_y_continuous(labels=scales::percent,
                     name="Observed - Expected \n Drought & Heat effect size") +
  scale_x_discrete(labels=~short_names_traits[.x], name=NULL) + 
  scale_fill_manual(values=c("Antagonistic"="#2A73AD","Additive"="white", "Synergistic"="#D98F07"),name="Effect") +
  scale_shape_manual(values=c("Antagonistic"=22,"Additive"=21, "Synergistic"=23),
                     name="Effect") + 
  geom_hline(yintercept=0, linetype="dashed", color="red"#, size=1.5
             ) 

# +
#   theme(
#     axis.title = element_text(size=80),
#     legend.title = element_text(size=80),
#     legend.text = element_text(size=80),
#     axis.text = element_text(size=70),
#     legend.key.size = unit(6,"line"),
#     legend.spacing = unit(3,"cm")
#   )
  


#ggsave("/Users/quino070/Bellwether_RGB/2025_paper/effect_by_trait.pdf", bg="transparent", height = 10)

```


```{r}
imap(resp_traits, ~{
  pt = summary_df %>% mutate(gen_cat = ifelse(
  genotype_name %in% c("CML333", "M162W", "Tzi8"), "Antagonistic", 
  ifelse(genotype_name %in% c("CML228", "CML322", "CML52","LH197"), 
         "Synergistic","Other"))) %>% 
    filter(day==13 & treatment!="Control" & gen_cat=="Antagonistic" & trait==.x) %>% 
    ggplot(aes(treatment, similarity)) + geom_point(color="darkgrey") + 
    geom_boxplot(color="#2A73AD", fill=NA
      ) +
    geom_line(aes(group=genotype_name), color="grey", linetype="dashed") + 
    facet_wrap(~trait, labeller = as_labeller(~short_names_traits[.x]), scales = "free") + 
    stat_compare_means(x="treatment", y="similarity", 
                                     facet.by="gen_cat", 
                                     comparisons = list(c("Drought", "Drought_Heat"),
                                          c("Drought_Heat","Heat"),
                                          c("Drought", "Heat")), method = "t.test",
                       aes(label = after_stat(p.signif))) +
  labs(x=NULL, y="Similarity to control") + scale_x_discrete(labels=treat_fullname) +
  theme_pubr() +
  theme(strip.background = element_rect(fill=NA), strip.text = element_text(size=12))
  
  pt = pt + scale_y_continuous(labels = scales::percent_format(), n.breaks = 4)
  
})   %>% ggarrange(plotlist = .)
  
#ggsave("/Users/quino070/Bellwether_RGB/2025_paper/ant_trends.pdf", height = 12, width = 10)
  


imap(resp_traits, ~{
  pt = summary_df %>% mutate(gen_cat = ifelse(
  genotype_name %in% c("CML333", "M162W", "Tzi8"), "Antagonistic", 
  ifelse(genotype_name %in% c("CML228", "CML322", "CML52","LH197"), 
         "Synergistic","Other"))) %>% 
    filter(day==13 & treatment!="Control" & gen_cat=="Synergistic" & trait==.x) %>% 
    ggplot(aes(treatment, similarity)) + geom_point(color="darkgrey") + 
    geom_boxplot(color="#D98F07", fill=NA
      ) +
    geom_line(aes(group=genotype_name), color="grey", linetype="dashed") + 
    facet_wrap(~trait, labeller = as_labeller(~short_names_traits[.x]), scales = "free") + 
    stat_compare_means(x="treatment", y="similarity", 
                                     facet.by="gen_cat", 
                                     comparisons = list(c("Drought", "Drought_Heat"),
                                          c("Drought_Heat","Heat"),
                                          c("Drought", "Heat")), method = "t.test",
                       aes(label = after_stat(p.signif))) +
  labs(x=NULL, y="Similarity to control") + scale_x_discrete(labels=treat_fullname) +
  theme_pubr() +
  theme(strip.background = element_rect(fill=NA), strip.text = element_text(size=12))
  
  pt = pt + scale_y_continuous(labels = scales::percent_format(), n.breaks = 4)
  
})   %>% ggarrange(plotlist = .)
  
#ggsave("/Users/quino070/Bellwether_RGB/2025_paper/syn_trends.pdf", height = 12, width = 10)

```
